### 2017-05-03 sheets/basis-1/sheet-basis-1.tex
basis1 ex1-3

\section{Diagrams}

\begin{hint}
  This is how you can draw diagrams:
  \[
    \begin{tikzcd}
                                    & A \arrow{dr}{g}    & \\
      B \arrow{ur}{f} \arrow{rr}{h} &                    & C
    \end{tikzcd}
  \]
\end{hint}

>>>>

%\section{Diagrams}
%
%\begin{hint}
%  This is how you can draw diagrams:
%  \[
%    \begin{tikzcd}
%                                    & A \arrow{dr}{g}    & \\
%      B \arrow{ur}{f} \arrow{rr}{h} &                    & C
%    \end{tikzcd}
%  \]
%\end{hint}

\section{Categories - Definition and Basics}

\begin {exercise}
Show there can be at most one inverse for a map $f \from A \to B$.
\end{exercise}

\begin {definition}{Rel}
The objects of the category \textbf{Rel} are sets, an arrow $f \from A \to B$ is a subset $f \sub A \times B$. 
The identity arrow on set $A$ is the equality relation $\{\angles{a,a} \such a \in A\}$. 
Composition of two maps $f \sub A \times B$, $g \sub B \times C$ is defined as 
\[ g \of f = \{ \angles{a, c} \in A \times C \such \exists b \emptybk (\angles {a, b} \in f \emptybk \& \emptybk \angles{b, c} \in g)\} \]
\end{definition}

\begin {exercise}
Show that \textbf{Rel} is a category.
\end{exercise}

\begin{definition}{Pos} The objects of  \textbf{Pos} are partially ordered sets 
(recall that a poset is a set $A$ equipped with a reflexive, transitive and antisymmetric binary relation $\leq^A_\cdot $ ) 
An arrow $ m \from A \to B $ function that is monotone, i.e. $\forall a, a' \in A$, 
\[a \leq^A_\cdot a' \implies m(a) \leq^A_\cdot m(a')\]
\end{definition}

\begin {exercise}
Show that \textbf{Pos} is a category.
\end{exercise}

### 2017-05-03 sheets/basis-1/sheet-basis-1s.tex
basis1 ex1-3

\begin {exercise}
Show there can be at most one inverse for a map $f \from A \to B$.
\end{exercise}

\begin {definition}{Rel}
The objects of the category \textbf{Rel} are sets, an arrow $f \from A \to B$ is a subset $f \sub A \times B$. 
The identity arrow on set $A$ is the equality relation $\{\angles{a,a} \such a \in A\}$. 
Composition of two maps $f \sub A \times B$, $g \sub B \times C$ is defined as 
\[ g \of f = \{ \angles{a, c} \in A \times C \such \exists b \emptybk (\angles {a, b} \in f \emptybk \& \emptybk \angles{b, c} \in g)\} \]
\end{definition}

\begin {exercise}
Show that \textbf{Rel} is a category.
\end{exercise}

@@

\begin {exercise}
Show that \textbf{Pos} is a category.
\end{exercise}

>>>>

\begin {exercise}
We have to show that for $f \from A \to B$, there is at most one $g \from B \to A$ s.t. $gf = 1_A$ and $fg = 1_B$. 
\\Assume that $g$ exists and that there is some $h \from B \to A$: $hf = 1_A$, $fh = 1_B$. 
\\We know from the identity axiom: $1_A \of g = g$ and $h \of 1_B = h$. Using this, our assumtions and the associativity axiom, we get:
 \[  1_A \of g = g \textiff (h \of f) \of g = g \textiff h \of (f \of g) = g \textiff h \of 1_B = g \textiff h=g. \]
\end{exercise}

\begin {definition}{Rel}
The objects of the category \textbf{Rel} are sets, an arrow $f \from A \to B$ is a subset $f \sub A \times B$. 
The identity arrow on set $A$ is the equality relation $\{\angles{a,a} \such a \in A\} := 1_A$. 
Composition of two maps $f \sub A \times B$, $g \sub B \times C$ is defined as 
\[ g \of f = \{ \angles{a, c} \in A \times C \such \exists b \emptybk (\angles {a, b} \in f \emptybk \& \emptybk \angles{b, c} \in g)\} \]
\end{definition}

\begin {exercise}
\textbf{identity laws:} for $f \sub A \times B$, 
\[f \of 1_A = \{ \angles{a, b} \in A \times B \such \exists a' \emptybk (\angles {a, a'} \in 1_A \emptybk \& \emptybk \angles{a', b} \in f)\} = \{ \angles{a, b} \in A \times B \such \exists a' \in A.  \emptybk \angles{a', b} \in f)\} = f.\] Analogously, $1_B \of f = f$.

\textbf{associativity:} for $f \sub A \times B, g\sub B \times C, h \sub C \times D$, 
\[
h \of (g \of f) = h \of \{ \angles{a, c} \in A \times C \such \exists b \emptybk (\angles {a, b} \in f \emptybk \& \emptybk \angles{b, c} \in g)\}
\]
\[= \{ \angles{a, d} \in A \times D \such \exists b, c \emptybk (\angles {a, b} \in f \emptybk \& \emptybk \angles{b, c} \in g \& \angles{c, d} \in h \} = (h \of g) \of f\]

\end{exercise}

@@

\begin {exercise}
\textbf{identity laws:} $1_A$ is the identity function, which is monotone since $a \leq^A_\cdot a' \implies a' \leq^A_\cdot a.$ 
\\The axiom holds because for $f \from A \in B, f \of 1_A = f = 1_B \of f$ (same as in \textbf{Set}).
\\
\textbf{associativity:} If $ f \from A \to B, g \from B \to C $ are monotone, so is $g \of f$ because \[a \leq^A_\cdot a' \implies f(a) \leq^A_\cdot f(a') \implies  g(f(a)) \leq^A_\cdot g(f(a'))  .\] Associativity follows from assiociativity of function composition.
\end{exercise}
